### 2011-09-15 doc/theses/mikula/analysis_of_xquery.tex
Master thesis: In progress.

\section{Structure of XML documents}
Most of queries can be exploited to obtain some information about a structure of respective XML documents. The structure of XML documents captures elements from these documents along with their names, attributes and their organization. What the \todo{Co to znamená? Root element?}outermost elements in these documents are, which elements can be contained within a certain element, whether they are optional or mandatory, etc.

@@

Additional path expressions \texttt{/bib/book/title}, \texttt{//author/name} indicate that element \texttt{book} also contains element \texttt{title} and element \texttt{author} contains element \texttt{name}. The latter one uses also descendant axis. \todo{A k čemu je tato informace při odvozování?}

@@

However, statements of this kind are necessary for the XML schema inference, their obtaining from queries is not significant due to the following reason. These statements could be determined directly from the XML documents and it could be easily done in a more convenient way. Queries may not cover the whole relevant content of the documents. For example some element may not be queried at all, therefore they may not be mentioned in any query \todo[inline]{Tady je ale několik případů - chtělo by popsat.
1) Element není zapsaný v dotazu, ale dostanu ho výsledku - např. rok anrození autora, ptám-li se na autory.
2) Element není zapsaný v dotazu, ale procházím ho při vyhodnocování -  např. to //autor
3) Element je v části dokumentu, která se dotazu vlbec netýká.}. In addition, these statements do not express any obligation of occurrence of elements and attributes or possible multiple occurrence of elements \todo{Tak toto je asi taky diskutabilní, např. // by mohlo naznačovat více výskytů.}. We cannot be sure that queries target nodes actually presented in the XML documents. Although query \texttt{/bib/book/author} indicates that element \texttt{author} is contained in element \texttt{book}, the query is valid whether this is true or not. In contrast, even basic methods of XML schema inference that utilize XML documents do not have these inadequacies.

@@

\texttt{lt} is a representant of so called \emph{value comparison operators} \todo{referencia} which operate two sequences of zero or one item. If an operand of a value comparison operator is a sequence of more than one item, then a \emph{type error} \todo{referencia} is raised.

@@

\subsection{XML Schema Built-in Types}
XML Schema built-in types of elements and attributes can be inferred from the XML documents by analysing their content. Since the count of built-in types is 44 and inheritance is involved, such analysis may be imprecise, especially when a large enough set of XML documents is not available. \todo[inline]{urcite priklad} This is when a set of XQuery queries may come in handy.

@@

Although determining built-in types of elements and attributes in simple queries may seem quite straightforward, reliable method would face some issues. More complex queries are commonly branched by \emph{if-then-else} construct. These branches can contain several contradictionary fragments from the view of type inference. Therefore, the reliable method would take the query structure into consideration and further research would be needed.

@@

\todo[inline]{Tady by to chtělo příklad. nebo to budeme řešit dál v navrženém algoritmu a toto je pouze diskuse případů?}

@@

\subsection{Approach from [TODO citace]}
Paper \todo{link} introduces a method of discovering keys and foreign keys by investigation of joins in queries. The basis of this discovery is a search for particular forms of joins, so called \emph{join patterns} \todo{uviest ich sem?}, but the joins are processed only if they are found in a particular syntactic form. Therefore joins with the same semantics written in different syntax are not taken into consideration.

@@

Apparently, element \texttt{part} can contain attributes \texttt{partid} and \texttt{partof}. Elements with unspecified attribute \texttt{partof} are at the top of the recursive hierarchy while, in elements with this attribute specified it makes a reference to an element with attribute \texttt{partid} of the same value. \todo{Tahle věta je nějaká divná.}

By using a similar approach as described in the paper \todo{ref} attribute \texttt{partid} can be marked as a key and attribute \texttt{partof} as its foreign key.

>>>>

\section{Structure of XML documents}
Most of queries can be exploited to obtain some information about a structure of respective XML documents. The structure of XML documents captures elements from these documents along with their names, attributes and their organization. What the root elements in these documents are, which elements can be contained within a certain element, whether they are optional or mandatory, etc.

@@

Additional path expressions \texttt{/bib/book/title}, \texttt{//author/name} indicate that element \texttt{book} also contains element \texttt{title} and element \texttt{author} contains element \texttt{name}. The latter one uses also descendant axis. That means the query considers all elements \texttt{author} in the document, thus it is a hint that maybe there are several elements named \texttt{author} but with different absolute paths and all of them contain an element \texttt{name}.

@@

However, statements of this kind are necessary for the XML schema inference, their obtaining from queries is not significant due to the following reason. These statements could be determined directly from the XML documents and it could be easily done in a more convenient way. Queries may not cover the whole relevant content of the documents. For example some elements may not be queried at all, therefore they may not be mentioned in any query \todo[inline]{Tady je ale několik případů - chtělo by popsat.
1) Element není zapsaný v dotazu, ale dostanu ho výsledku - např. rok anrození autora, ptám-li se na autory.
2) Element není zapsaný v dotazu, ale procházím ho při vyhodnocování -  např. to //autor
3) Element je v části dokumentu, která se dotazu vlbec netýká.}. In addition, these statements do not express any obligation of occurrence of elements and attributes nor clearly determine multiple occurrence of elements. We cannot be sure that queries target nodes actually presented in the XML documents. Although query \texttt{/bib/book/author} indicates that element \texttt{author} is contained in element \texttt{book}, the query is valid whether this is true or not. In contrast, even basic methods of XML schema inference that utilize XML documents do not have these inadequacies.

@@

\texttt{lt} is a representant of so called \emph{value comparison operators} \todo{referencia na def} which operate two sequences of zero or one item. If an operand of a value comparison operator is a sequence of more than one item, then a \emph{type error} \todo{referencia na def} is raised.

@@

\subsection{XML Schema Built-in Types}
XML Schema built-in types of elements and attributes can be inferred from the XML documents by analysing their content. Since the count of built-in types is 44 and inheritance is involved, such analysis may be imprecise, especially when a large enough set of XML documents is not available. This is when a set of XQuery queries may come in handy. For example, consider the following occurrences of element \texttt{a}.

\begin{verbatim}
<a>1</a>
<a>6</a>
<a>18</a>
\end{verbatim}

These three occurrences are not sufficient enough to determine the accurate type of element \texttt{a}, because values \texttt{1}, \texttt{6}, \texttt{18} are valid values of several types: \texttt{decimal}, \texttt{integer}, \texttt{byte}, \texttt{short}, \texttt{unsignedInt}, \texttt{positiveInteger} and many others. The following text introduces some possibilities of a type inference from queries.

@@

Although determining built-in types of elements and attributes in simple queries may seem quite straightforward, reliable method would face some issues. More complex queries are commonly branched by \emph{if-then-else} construct. These branches can contain several contradictionary fragments from the view of type inference. Therefore, the reliable method would take the query structure into consideration and further research would be needed.
\todo[inline]{Tady by to chtělo příklad. nebo to budeme řešit dál v navrženém algoritmu a toto je pouze diskuse případů?}

@@

@@

\subsection{Approach from \cite{Necasky:2009:DXK:1529282.1529414}}
Paper \todo{link} introduces a method of discovering keys and foreign keys by investigation of joins in queries. The basis of this discovery is a search for particular forms of joins, so called \emph{join patterns} \todo{uviest ich sem?}, but the joins are processed only if they are found in a particular syntactic form. Therefore joins with the same semantics written in different syntax are not taken into consideration.

@@

Apparently, element \texttt{part} can contain attributes \texttt{partid} and \texttt{partof}. Elements with unspecified attribute \texttt{partof} are at the top of the recursive hierarchy while, each element with this attribute specified references an element with attribute \texttt{partid} of the same value.

By using a similar approach as described in \cite{Necasky:2009:DXK:1529282.1529414} attribute \texttt{partid} can be marked as a key and attribute \texttt{partof} as its foreign key.
